Mergers in the COSMOS Field

### 2023-11-29 paper-source/main.tex
Major update to finding weighted AGN fraction. Also added to paper and lots of projected separation stuff. Paper nearly there, just gotta write a discussion and conclusion.

To better our understanding of galaxy interaction, it is often prudent to break them down into different stages of interaction. This relates to the dynamical time of the interaction, and the snapshot in the interaction we are observing. Previous works focused on simulations find that a sudden increase in star formation occurs early in the dynamical timescale of the interaction \citep{2008MNRAS.384..386C, 2019MNRAS.490.2139R}, however, confirming this definitive link observationally remains elusive \citep{2023arXiv230916531R}. The same studies could also be done for nuclear activation as well as quenching of interacting systems \citep{Papers on nuclear activation}. While we cannot yet directly measure when in the dynamical timescale star formation is enhanced from observations, we can approximate it by splitting different interactions into different stages. These stages relate directly to the dynamical timescale of the interaction. Often, they are split into four: gravitationally bounded pairs, approaching pericentre of the interaction, approaching apocentre of the interaction, and approaching coalescence.

@@

This paper is laid out as follows in Section \ref{data} we explain the OR23 and COSMOS catalogues, and describe the process of catalogue matching especially focused on accurate de-duplication and cross-matching selection from the different catalogues.  Section \ref{staging} describes the methods by which we split the interacting systems into their different stages, and how we identify the secondary galaxies to each identified interacting galaxy. We also discuss our processes of calculating projected separations. In Section \ref{results} we describe and discuss our results, and their implications, and compare to previous works focused on investigating different stages of interaction. Finally, Section \ref{conclusion} we make concluding remarks and discuss future work to better our constraints.

@@

\subsection{Secondary Identification}
\noindent As stated previously, the OR23 catalogue only identified interacting galaxy systems, not individual interacting galaxies. Therefore, we must find the secondary in each interaction using the COSMOS2020 catalogue. The size of the cutouts in this catalogue were also small, and therefore, the secondary did not often appear in the cutout in that project. To find the secondaries, we apply three steps. First, a cutout surrounding the primary of each system was created. These cutouts were from the COSMOS cutout service, selecting HST-ACS tiles as the source of the cutouts in the $F814W$ filter. Each cutout had a 30" radius (corresponding to 1001 $\times$ 1001 pixels) centred and was centred on the primary galaxy. The original OR23 cutout was also displayed next to the enlarged cutout. We then searched the COSMOS2020 catalogue for all sources within the cutout. Their positions and redshifts were recorded and used to annotate the cutout.

We then visually assessed each annotated cutout, giving one of the four following classifications to each: system disturbed but secondary could not be identified; secondary could be identified; cannot confirm galaxy is interacting; null redshift; incorrect primary assigned. Our definition of a secondary galaxy was such that the secondary had to be within the cutout we were visually assessing and within $z=\pm0.1$ of the primary redshift. We use this cutoff as this is often used when calculating environment parameters \citep[e.g][]{Baldry 2006} or defining interacting galaxies by close pairs \citep[e.g][]{Shah et al. paper from 2022 / 2023} based on the uncertainties present in using photometric redshift estimates from SED fitting. We define a null redshift as one outwith our redshift limits or NaN. A minority of the cutouts we visually assessed were found to have the incorrect primary at the centre. This was a result of our cross matching based on radius with the COSMOS catalogue.

@@

@@

\section{METHOD: Environment, AGN and Interaction Stage}
In this section, we describe how we separate our sample of 4,181 interacting galaxies into separate stages. Each stage is defined to capture a different part of the interaction, and a different part of the dynamical history of the interaction. We want to compare different stages in the interaction to different physical underlying processes, such as star formation rate. Such parameters can be found in the COSMOS2020 catalogue. However, two fundamental properties we also want to compare with is that of AGN fraction evolution with time and that of environment, and the impacts this might have.

@@

@@

Stage 3 interacting galaxies are described as those interacting systems where the systems have approached apocentre from each other. In many of cases of this stage, if the galaxies have sufficient velocity, the galaxies will escape each other and not coalesce. Hence, galaxies at this stage often could not have a secondary galaxy identified in our visual assessment. We purely describe a galaxy as stage 3 based on significant tidal and morphological disturbance. This stage has the highest potential for contamination by irregular galaxies or galaxies whose morphology has been disturbed by processes other than tidal interaction. We discuss the potential impacts of this contamination, and the limitations of using visual morphology as a proxy for stage in Section \ref{sec:limitations}.

@@

Our four stage approach is not a new one, and many other works have utilised it to differentiate different parts of the dynamical history of a galaxy interaction \citep{}. It is, however, important to note its limitations. First, this is a simplified view of the dynamical timescale of an interaction. We have assumed that an interaction begins, the pair have enough velocity to escape or they coalesce and that is representative of the full dynamical timescale. We have not considered systems where there are multiple flybys before coalescence or systems where there are multiple members at different stages of the interaction. This is a particular concern when looking at stages 2 and 3. For instance, the effects could be enhanced or diminished over multiple flybys, and we have no way to capture this.

However, with these simplifications in our assumptions, we will investigate the parameters which are proxies for the underlying physical processes and interpret them. We will investigate the whether we see have a bias in our systems based on environment, whether we find increased AGN activity and if we observe enhancement in star formation with stage. The star formation rate is already present in the COSMOS2020 catalogue from both EAzY and LePhare. For the environment parameter and whether AGN are detected, we cross match to two further catalogues. We describe these catalogues here, our process of cross matching and the new sample sizes we have in each subsample.

@@

\subsection{Classifying AGN}\label{sec:agn-clsf}
\noindent We will also investigate the affect of interaction stage upon the AGN activity within the galaxies involved. Unfortunately, the outputs of EAzY and LePhare do not provide us with photometric estimates of the line fluxes needed to define an AGN on a Baldwin, Philips and Terlevich (BPT) diagram \citep{}. Therefore, we cannot directly identify if a galaxy is only star forming, has nuclear activity or is a composite. So, we turn to cross-matching our sample to three different catalogues which contain the required information. We use the Chandra COSMOS Legacy Survey Multiwavelength Catalogue \citep{2016ApJ...817...34M} and the COSMOS VLA 3GHz survey \citep{2017A&A...602A...6S, 2017A&A...602A...3D}. Each source has an AGN, star forming or composite classification is within the catalogue.

@@

While the ratio of AGN to star forming galaxies in our sample seems large compared to other works, it is important to note that this is a result of limited matching between the catalogues. Of the 4,182 galaxies in our sample, only 1,214 appeared at all in either the VLA or Chandra catalogues. Any galaxy which did not appear in these catalogues, we discard as unclassified. Therefore, we require these catalogues to specifically show us if a galaxy is star forming or not before we classify it as such. This is why our number of star forming galaxies is lower than expected. We discuss this limitation in section \ref{sec:limitations} and the effect it may have on our results of AGN fraction evolution with stage.

@@

Figure \ref{fig:weighted-mass} shows the weighted mass distributions through the four different stages. These appear highly similar, and we conduct Kolmogorov-Smirnov \citep[KS-test;][]{an1933sulla} and Anderson-Darling \citep[AD-test;][]{AD_paper} tests to compare each distribution. In each test, we calculate the value of the KS test and a p-value which is the probability that each distribution is drawn from the same parent distribution. For each mass distribution, we find that p-value of the KS test is approximately one. Thus, it is not unreasonable to assume that the galaxies between each of our samples are drawn from the same parent sample and are very similar. Figure \ref{fig:weighted-sfr} shows the SFR distributions while being weighted and controlled for mass. When comparing stages 1 and 2, 1 and 3, 1 and 4, 2 and 4 and 3 and 4, the p-values are $\ll$0.05. Thus, we can reject the null hypothesis for these distributions and assume they are from different parent samples and are not identical. However, for comparing stages 2 and 3, the p-value of the KS test is $\approx$0.90. Thus, while these distributions are likely to be not identical, they are very similar in the parent sample they have been drawn from. \textbf{Thus, for the same mass distribution through each stage of interaction, the star formation distribution radically changes from stage 1 to stage 2 and from stage 3 to stage 4, while remaining reasonably similar from stage 2 to stage 3.}

@@

% So, what does this all mean?
Thus, we find that throughout the different stages of interaction the star formation within the systems is being driven upwards. This is not an unexpected result. It has long been known that the SFRs of merging and post-merging galaxies is increased with respect to control samples of galaxies within the field \citep{papers saying this}. However, what we have found here is that this process of increasing SFR and galaxies being driven to starburst begins from the pericentre of the first interaction. Simulations have long showed that there is a peak in SFR initially in an interaction, followed by a slow reduction to a final enhanced level in the post-merger stage \citep{papers saying this}. However, due to the initial starburst, it is often shown that the gas within the galaxy is either quickly used up and the galaxy quenches or it is heated to the point that star formation becomes very inefficient. Thus, what we would expect to see here is that stages 2 and 4 would have the highest fractions of galaxies undergoing a starburst and the smallest quiescent fractions.

@@

Now, it is important to note the parameter space that we are searching in these examples. We are probing major interactions between galaxies, where the resultant tidal features that form would be classifiable in an image. It is well know that the mass ratio of an interaction takes an important role in the enhancement of star formation \citep{papers about major interaction}, and we are probing the most extreme examples. Hence, even those mergers which are considered dry - with very little gas - would be capable of driving major rejuvenation in the star formation of these systems. To conclude, \textbf{we find clear, statistically robust evolution of a galaxys' SFR with interaction stage. The SFR changes dramatically from stage 1 to 2 and stage 3 to 4 - at the pericentre and coalescence of the interaction. Existing simulations of galaxy interactions support this interpretation of when star formation changes.}

\subsection{Controlling for Galactic Environment} \label{sec:env-cont}
\noindent While we have found clear evidence of evolution in the SFR of interacting galaxies with stage, it is important we remove a potential source of contamination. It is well known that the environment has a direct impact on the observed SFR of galaxies. A galaxy in a cluster environment has, on average, a higher SFR than those in the field \citep{Papers on environment with SFR}. Thus, if any of our stage classifications are biased towards one environment or another, it could have severe consequences for our results. Therefore, we utilise the existing COSMOS environment catalogue described in \citet{2017ApJ...837...16D} to ensure there are no environmental biases in our staging.

@@

\section{Nuclear Activity with Interaction Stage}
\subsection{Controlling for Stage}
\noindent We now cross match our sample with the two AGN catalogues and methods described in Section \ref{sec:agn-clsf}. This gives a sample of 1,284 AGN and star forming galaxies (SFGs) to cross match with the stage. Figure \ref{fig:agn-stage} shows distribution of stellar mass to SFR with stage, with the confirmed AGN and SFG marked.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/agn-stage-dist.pdf}
    \caption{The distribution of AGN through stage with SFR and stellar mass.}
    \label{fig:agn-stage}
\end{figure}

\section{Characteristics of Matched Pairs}
\subsection{Star Formation and Mass Ratio}

\subsection{Star Formation and Projected Separation}

\section{Limitations of Approach}\label{sec:limitations}
\subsection{Redshift Effects on Visual Classification}
\noindent An important limitation of our visual approach is related to the effect of tidal features out to higher redshifts. As the redshift increases, surface brightness dimming of the systems may lead to a lack of visual confirmation in the tidal features of different objects. Thus, if we were heavily affected by this in our redshift range ($0.0 \leq z \leq 1.2$), we would see distinct signs of it in our selection function of our mergers. For instance, if all of our stage 4 classifications were actually high redshift stage 3 galaxies, we would observe this in our selection function. Therefore, Figure \ref{fig:redshift_selection} shows our selection functions across our redshift range for each defined stage in our interaction.

As shown, the distribution of masses across redshift that we find in our sample is reasonably consistent with each other. This means we are likely not affected by surface brightness dimming across our sample and our redshift range.

\subsection{Selection Effects in AGN Sample}

\subsection{Limitations of Visual Classification}

\subsection{Selection Effects of SFR Sample}

>>>>

To better our understanding of galaxy interaction, it is often prudent to break them down into different stages of interaction. This relates to the dynamical time of the interaction, and the snapshot in the interaction we are observing. Previous works focused on simulations find that a sudden increase in star formation occurs early in the dynamical timescale of the interaction \citep{2008MNRAS.384..386C, 2019MNRAS.490.2139R}, however, confirming this definitive link observationally remains elusive \citep{2023arXiv230916531R}. The same studies could also be done for nuclear activation as well as quenching of interacting systems \citep{2011MNRAS.418.2043E, 2018PASJ...70S..37G, 2023ApJ...942..107S}. While we cannot yet directly measure when in the dynamical timescale star formation is enhanced from observations, we can approximate it by splitting different interactions into different stages. These stages relate directly to the dynamical timescale of the interaction. Often, they are split into four: gravitationally bounded pairs, approaching pericentre of the interaction, approaching apocentre of the interaction, and approaching coalescence.

@@

This paper is laid out as follows in Section \ref{data} we explain the OR23 and COSMOS catalogues, and describe the process of catalogue matching especially focused on accurate de-duplication and cross-matching selection from the different catalogues.  Section \ref{method} describes the methods by which we split the interacting systems into their different stages, and how we identify the secondary galaxies to each identified interacting galaxy. We also discuss our processes of calculating projected separations. In Section \ref{results} we describe and discuss our results, and their implications, and compare to previous works focused on investigating different stages of interaction. Finally, Section \ref{conclusion} we make concluding remarks and discuss future work to better our constraints.

@@

\subsection{Secondary Identification}\label{sec:sec-ident}
\noindent As stated previously, the OR23 catalogue only identified interacting galaxy systems, not individual interacting galaxies. Therefore, we must find the secondary in each interaction using the COSMOS2020 catalogue. The size of the cutouts in this catalogue were also small, and therefore, the secondary did not often appear in the cutout in that project. To find the secondaries, we apply three steps. First, a cutout surrounding the primary of each system was created. These cutouts were from the COSMOS cutout service, selecting HST-ACS tiles as the source of the cutouts in the $F814W$ filter. Each cutout had a 30" radius (corresponding to 1001 $\times$ 1001 pixels) centred and was centred on the primary galaxy. The original OR23 cutout was also displayed next to the enlarged cutout. We then searched the COSMOS2020 catalogue for all sources within the cutout. Their positions and redshifts were recorded and used to annotate the cutout.

We then visually assessed each annotated cutout, giving one of the four following classifications to each: system disturbed but secondary could not be identified; secondary could be identified; cannot confirm galaxy is interacting; null redshift; incorrect primary assigned. Our definition of a secondary galaxy was such that the secondary had to be within the cutout we were visually assessing and within $z=\pm0.1$ of the primary redshift. We use this cutoff as this is often used when calculating environment parameters \citep[e.g][]{2006MNRAS.373..469B} or defining interacting galaxies by close pairs \citep[e.g][]{2022ApJ...940....4S} based on the uncertainties present in using photometric redshift estimates from SED fitting. We define a null redshift as one outwith our redshift limits or NaN. A minority of the cutouts we visually assessed were found to have the incorrect primary at the centre. This was a result of our cross matching based on radius with the COSMOS catalogue.

@@

Each of the galaxies that have a confirmed secondary will be used in measuring effects of projected separation with star formation and AGN fraction. Therefore, we must define a control sample by which to compare to. We find these control galaxies from the COSMOS2020 catalogue. All galaxies within 0.01 dex of our sub-samples' stellar mass are selected from the catalogue, and within a $\pm$0.01 redshift slice. We then order the control galaxies to be whichever system is furthest from our target. Each control galaxy is then visually confirmed to be non-interacting and in isolation. We also confirm that it does not already exist in the catalogue, and ensure there is no duplication in the control sample. Figure \ref{fig:matched-distributions} shows the mass distribution of our paired sub-sample of galaxies and their control, showing a closely similar mass distribution. With the previously defined cuts, we find a control galaxy for all but one of our galaxies with a confirmed secondary.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/mass-matching-pairs.pdf}
    \caption{The mass distribution of the paired interacting galaxy sample and the control sample. Both primary and secondary galaxies are within this distribution. From our sample of 4,135 interacting galaxies morphologically identified, 960 are confirmed galaxy pairs. This means that the secondary in the interacting galaxy system has been identified from morphology and photometric redshifts.}
    \label{fig:matched-distributions}
\end{figure}

@@

\subsection{Redshift Effects on Visual Classification}
\noindent There are two limits when using visual morphology to find interacting galaxies. The first is the limitation of being able to characterise tidal features with increasing redshift. As the redshift increases, surface brightness dimming of the systems may lead to a lack of visual confirmation in the tidal features of different objects. Thus, if we are heavily affected by this in our redshift range ($0.0 \leq z \leq 1.2$), we would see distinct signs of it in our selection function of our mergers. For instance, if all of our stage 4 classifications were actually high redshift stage 3 galaxies, we would observe this in our selection function. Therefore, Figure \ref{fig:redshift_selection} shows our selection functions across our redshift range for each defined stage in our interaction. As shown, the distribution of masses across redshift that we find in our sample is reasonably consistent with each other. This means we are likely not affected by surface brightness dimming across our sample and our redshift range.

The second issue that could affect our sample is that of contamination by close pairs by projection effects rather than by galaxies which are physically close together. For the majority of these systems, we are limited to only having their calculated photometric redshift from the COSMSO2020 catalogue and not their well-constrained spectroscopic redshift. As we are conducting visual classification, this issue is rather limited. The criteria on all of our sources besides stage 1 is that they must be morphologically disturbed. Each source is confirmed to have some kind of tidal feature f=present in the image, even if a secondary is not present in the image. There are, always, of course other galaxies in each image and there is no guarantee that the calcualted photometric redshift is correct. WE often find galaxies which are clearly interacting or paired, but with the photometric wildly different. In these cases, we defer to the redshift and only record one of the galaxies in our sample. This approach, however, means we must be careful to investigate the effects of environment on our galaxies as disturbance by, say, ram pressure stripping can appear like tidal features.

\section{METHOD: Environment, AGN and Interaction Stage} \label{method}
In this section, we describe how we separate our sample of 4,181 interacting galaxies into separate stages. Each stage is defined to capture a different part of the interaction, and a different part of the dynamical history of the interaction. We want to compare different stages in the interaction to different physical underlying processes, such as star formation rate. Such parameters can be found in the COSMOS2020 catalogue. However, two fundamental properties we also want to compare with is that of AGN fraction evolution with time and that of environment, and the impacts this might have.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.95\textwidth]{figures/projected-seps.pdf}
    \caption{The projected separations of the confirmed galaxy pairs in our sample. This confers with other works the definition of our different stages. Stage 1 can be at any projected separation, however, we have visually confirmed that these sources are not morphologically disturbed. Stage 2 is dominated by systems with small projected separation as, by definition, they must be morphologically linked or overlapping. Those systems at larger separation are very large systems whose morphology categorise them as stage 2. Finally, stage 3 galaxies are those which are visually confirmed to be fully morphologically separated and tidally disturbed. The bulk of these lie in a range of 25 - 100kpc in projected separation from each other. There is some overlap between stage 2 and stage 3 in projected separation, as their visual classification is also dependent on the system size.}
    \label{fig:proj-seps}
\end{figure*}

@@

Stage 3 interacting galaxies are described as those interacting systems where the systems have approached apocentre from each other. In many of cases of this stage, if the galaxies have sufficient velocity, the galaxies will escape each other and not coalesce. Hence, galaxies at this stage often could not have a secondary galaxy identified in our visual assessment. We purely describe a galaxy as stage 3 based on significant tidal and morphological disturbance.

@@

Our four stage approach is not a new one, and many other works have utilised it to differentiate different parts of the dynamical history of a galaxy interaction \citep{2022ApJ...937...97C, 2023ApJ...952..122G}. It is, however, important to note its limitations. First, this is a simplified view of the dynamical timescale of an interaction. We have assumed that an interaction begins, the pair have enough velocity to escape or they coalesce and that is representative of the full dynamical timescale. We have not considered systems where there are multiple flybys before coalescence or systems where there are multiple members at different stages of the interaction. This is a particular concern when looking at stages 2 and 3. For instance, the effects could be enhanced or diminished over multiple flybys, and we have no way to capture this.

We relate our classification of stage to projected separations between systems, more often used in the literature, by using our confirmed galaxy pair subsample. Figure \ref{fig:proj-seps} shows the stage classification with projected separation between the two pairs. The projected separation is measured by taking the average of the best fit photometric redshifts between the two galaxies, and converting their angular separation to a physical one. The most dinstinct separation in stages in between stage 2 and stage 3. Here, we see that stage 2 is dominated by systems with a projected separation $<$35kpc while stage 3 is dominated $25 \leq s_{proj} \leq 100$ kpc. This is to be expected from our definitions of stage. Stage 2 is visually classified as systems which are highly morphologically disturbed, while either being morphologically linked to each other or overlapping. Those galaxy pairs with large projected separations are pairs which are very large in angular size, while still overlapping or morphologically linked. If this criteria is not met, then the system becomes a stage 3 interaction where the two galaxies in the pair are completely dinstinct. There is some overlap between the projected separations of stage 2 and stage 3 as this is somewhat dependent on the angular size of the two systems involved in the interaction.

We will investigate the parameters which are proxies for the underlying physical processes of our sample and interpret them. We will investigate whether we see have a bias in our systems based on environment, whether we find increased AGN activity and if we observe enhancement in star formation with stage. The star formation rate is already present in the COSMOS2020 catalogue from both EAzY and LePhare. For the environment parameter and whether AGN are detected, we cross match to two further catalogues. We describe these catalogues here, our process of cross matching and the new sample sizes we have in each subsample.

@@

\subsection{Classifying AGN}\label{sec:agn-clsf}
\noindent We will also investigate the affect of interaction stage upon the AGN activity within the galaxies involved. Unfortunately, the outputs of EAzY and LePhare do not provide us with photometric estimates of the line fluxes needed to define an AGN on a Baldwin, Philips and Terlevich (BPT) diagram. Therefore, we cannot directly identify if a galaxy is only star forming, has nuclear activity or is a composite. So, we turn to cross-matching our sample to three different catalogues which contain the required information. We use the Chandra COSMOS Legacy Survey Multiwavelength Catalogue \citep{2016ApJ...817...34M} and the COSMOS VLA 3GHz survey \citep{2017A&A...602A...6S, 2017A&A...602A...3D}. Each source has an AGN, star forming or composite classification is within the catalogue.

@@

While the ratio of AGN to star forming galaxies in our sample seems large compared to other works, it is important to note that this is a result of limited matching between the catalogues. Of the 4,182 galaxies in our sample, only 1,214 appeared at all in either the VLA or Chandra catalogues. Any galaxy which did not appear in these catalogues, we discard as unclassified. Therefore, we require these catalogues to specifically show us if a galaxy is star forming or not before we classify it as such.

@@

Figure \ref{fig:weighted-mass} shows the weighted mass distributions through the four different stages. These appear highly similar, and we conduct Kolmogorov-Smirnov \citep[KS-test;][]{an1933sulla} and Anderson-Darling \citep[AD-test;][]{stephens_74} tests to compare each distribution. In each test, we calculate the value of the KS test and a p-value which is the probability that each distribution is drawn from the same parent distribution. For each mass distribution, we find that p-value of the KS test is approximately one. Thus, it is not unreasonable to assume that the galaxies between each of our samples are drawn from the same parent sample and are very similar. Figure \ref{fig:weighted-sfr} shows the SFR distributions while being weighted and controlled for mass. When comparing stages 1 and 2, 1 and 3, 1 and 4, 2 and 4 and 3 and 4, the p-values are $\ll$0.05. Thus, we can reject the null hypothesis for these distributions and assume they are from different parent samples and are not identical. However, for comparing stages 2 and 3, the p-value of the KS test is $\approx$0.90. Thus, while these distributions are likely to be not identical, they are very similar in the parent sample they have been drawn from. \textbf{Thus, for the same mass distribution through each stage of interaction, the star formation distribution radically changes from stage 1 to stage 2 and from stage 3 to stage 4, while remaining reasonably similar from stage 2 to stage 3.}

@@

% So, what does this all mean?
Thus, we find that throughout the different stages of interaction the star formation within the systems is being driven upwards. We have found here is that this process of increasing SFR and galaxies being driven to starburst begins from the pericentre of the first interaction. Simulations have long showed that there is a peak in SFR initially in an interaction, followed by a slow reduction to a final enhanced level in the post-merger stage. However, due to the initial starburst, it is often shown that the gas within the galaxy is either quickly used up and the galaxy quenches or it is heated to the point that star formation becomes very inefficient. Thus, what we would expect to see here is that stages 2 and 4 would have the highest fractions of galaxies undergoing a starburst and the smallest quiescent fractions.

@@

Now, it is important to note the parameter space that we are searching in these examples. We are probing major interactions between galaxies, where the resultant tidal features that form would be classifiable in an image. Hence, even those mergers which are considered dry - with very little gas - would be capable of driving major rejuvenation in the star formation of these systems. To conclude, \textbf{we find clear, statistically robust evolution of a galaxys' SFR with interaction stage. The SFR changes dramatically from stage 1 to 2 and stage 3 to 4 - at the pericentre and coalescence of the interaction. Existing simulations of galaxy interactions support this interpretation of when star formation changes.}

\subsection{Projected Separation and Star Formation Enhancement}
\noindent To further confirm this result, we directly investigate the relation between the star formation enhancement (SFE) and the projected separation using our confirmed sub-sample of galaxy pairs. This sample is significantly smaller than our non-pair sample: containing 480 pairs or 960 galaxies. Upon sub-dividing this into different stages we find 310 stage 1 galaxies, 146 stage 2 galaxies and 498 stage 3 galaxies. Only 3 stage 4 galaxies were in our confirmed galaxy pair sub-sample, and therefore we do not attempt to make inferences about this population.

To measure the SFE of our galaxy pairs, we directly compare the to the control sample that was also created with this sub-sample and defined in Section \ref{sec:sec-ident}. We separate our galaxy pair sample into different bins based on the projected separation between them. We find that the bulk of our sample has a projected separation between the two galaxies of $\leq$50kpc. Therefore, we sample from this region of the parameter space with high precision and smaller bin widths before we increase the bin widths at larger projected separations. We define a cutoff that each bin must contain at least 10 counts to be included in this plot and, therefore, by changing the bin widths with projected separation we are able to maintain some level of statistical robustness. We define our bins as [0.5, 10, 20, 50, 100, 125, 150] kpc.

We create distributions based on these bins of projected separation: one of interacting galaxies and the other of our control galaxies. We then take the average of each bin separately, before dividing each distribution by the other. By taking the average of each bin, we are finding the total excess of the SFR caused by interaction alone and then comparing it to what we would expect from non-interacting galaxies in the same bin. Note, that the control galaxy relation to the projected separation is meaningless as they are not paired. The control galaxies are simply mass and redshift matched to each interacting galaxy and used as a reference for what SFR we would expect of it had it not been interacting. Finally, we divide the the averaged interacting SFR distribution by the averaged control SFR distribution. This provides us with a measure of the excess star formation due to interaction, and the SFE when compared to the regular population.

\begin{figure*}
    \centering
    \includegraphics[width=0.95\textwidth]{figures/sfr-enhancement-projected-sep.pdf}
    \caption{The projected separation against the star formation enhancement in average star formation at different bins of projected separation. Each bin must contain at least 10 counts to be considered. As the bulk of our galaxy pair sample is at low projected separation, we heavily sample from this region of the parameter space. The bins are: [0.5, 10, 20, 50, 100, 125] kpc. As we move to higher projected separation, the bins increase in width to maintain statistical significance in our sample. \textit{Left}: The star formation enhancement found in the entire galaxy pair sample. As expected, we see a gradually decreasing enhancement. \textit{Right}: As left but broken up into different stages of the interaction. Black markers are stage 1, red stage 2 and blue stage 3. We only investigate stages 1,2 and 3 as only three galaxy pairs were identified in stage 4. Here, we find a generally decreasing star formation enhancement with projected separation but very different individual behaviour dependent on the stage classification.}
    \label{fig:sfr-enhancement-sep}
\end{figure*}

Figure \ref{fig:sfr-enhancement-sep} shows star formation enhancement between our interacting and control binned SFRs and the projected separations between each galaxy pair. On the left of the plot, we have the distribution for the full galaxy pair sample, without taking account of stage. The error bars are calculated based on the counts within the bin and errors on the underlying SFRs of the galaxies within. The dashed black line represents no enhancement in SFR, as the average SFR in the interacting galaxy sample would be equal to the average SFR in the control sample. We find at very small projected separation a SFR enhancement of 1.87 which gradually decreases with projected separation down to approximately 1 at 117.5kpc. On the right, we break the galaxy pair sample into different stages and plot out the resultant enhancement in star formation. This shows very different behaviour in enhancement dependent on the stage of the interaction.

In stage 1, we find that the star formation is very weakly enhanced, at a level of approximately 1.25. As we increase the projected separation bins, the enhancement then changes around an enhancement of 1, even dropping below 1 at 40kpc. Taking into account the errors on our measurements, the true value of the enhancement could be around 1 for the entire projected separation distribution. This is not unexpected. Stage 1 represents when the two galaxies are close pairs, but with no morphological disturbance. Therefore, we would not expect significant, if any, enhancement in the SFR based on the projected separation. This is the stage with the lowest enhancement across projected separation, remaining close to 1.

In stage 2, we see consistent enhancement with projected separation. It is also much larger than stage 1, with it being approximately 1.75 up to 1.90. In fact, the measured enhancement increases across the 50kpc of projected separation. However, accounting for the errors on these measurements and the declining counts in each bin, it is likely that the enhancement does drop as the projected separation increases. We see a very large enhancement in the overlapping stage 3 galaxies which overlap with the stage 2 measurements. The enhancement in stage 3 galaxies then rapidly declines as we move to higher projected separations. After 150kpc, our galaxy pair sample does not have the counts to make robust estimates of the SFE.

Finding that the SFE declines with projected separation is not an unexpected. Previous works also find this and even match our results \citep{2000ApJ...530..660B, 2008AJ....135.1877E, 2013MNRAS.433L..59P}. Where our result differs is by taking into account the stages of the interaction. \textbf{While using only projected separation is a proxy for stage, we find the actual SFE is also dependent on the time of in the dynamical timescale the interaction is}. For instance, all of our stage 2 systems have high SFE of greater than 1.5 no matter the projected separation between them. This shows that the morphological connection and overlap of the two systems is driving the enhancement. Then, moving to stage 3 with low projected separation, we see the SFE grows even larger to 2.5. The galaxy pair will be just at the point of breaking overlap and morphological connection here, only just moving away from pericentre here. Simulations have predicted that once the two systems are just passed pericentre and moving away from each other, the two galaxies will undergo an intense delayed starburst \citep{2015MNRAS.448.1107M}. We see this effect in the sudden increase in SFE as the two galaxies disconnect from each other and begin to move away from one another.

These distributions of SFE paint a picture of the disturbance and change in the SFRs of these systems throughout the interaction, we will summarise here. First, in stage 1, the two systems move towards one another (or away with little disturbance) with no SFE and matching the control. Upon entering into stage 2, the galaxies SFR dramatically increases due to the morphological disturbance of the two disks and potential overlapping and linking of the two systems. Then, once the two systems are past pericentre and enter into stage 3 to flyby one another, the gravitational torques on the two systems take hold and star formation is, again, dramatically increased. This increase relaxes as the systems move towards apocentre and higher projected separations.

However, it is important to note that this result has been found with a sample of 960 interacting galaxies in their pairs. No stage 4 galaxies have been represented and in numerous projected separation bins the counts are small enough to have conclusion altering error bars. \textbf{Therefore, it is imperative that future works, with larger sample sizes not only look at the projected separation of the systems but the morphology as well.}

\subsection{Controlling for Galactic Environment} \label{sec:env-cont}
\noindent While we have found clear evidence of evolution in the SFR of interacting galaxies with stage, it is important we remove a potential source of contamination. It is well known that the environment has a direct impact on the observed SFR of galaxies. A galaxy in a cluster environment has, on average, a higher SFR than those in the field \citep{2006MNRAS.373..469B}. Thus, if any of our stage classifications are biased towards one environment or another, it could have severe consequences for our results. Therefore, we utilise the existing COSMOS environment catalogue described in \citet{2017ApJ...837...16D} to ensure there are no environmental biases in our staging.

@@

\section{Nuclear Activity with Interaction Stage}
\noindent We now cross match our sample with the two AGN catalogues and methods described in Section \ref{sec:agn-clsf}. This provides us a sample of 1,361 AGN and star forming galaxies (SFGs). The breakdown of number of classifications per stage is shown in Table \ref{tab:agn-sfg-breakdown}. Figure \ref{fig:agn-stage} shows distribution of stellar mass to SFR with stage, with the confirmed AGN and SFGs marked. We find no major changes in AGN with stellar mass or SFR with stage in our sample. Galaxies containing AGN appearing throughout the SFMS and red sequence across each stage. There is no obvious bias in stellar mass or SFR for hosting an AGN. We do, however, see a concentration of AGN in stage 1 on the red sequence at higher masses. This shifts as we increase interaction stage, with the bulk of the AGN beginning to appear at lower masses in stages 3 and 4. Finally, in stage 4 we see AGN and SFGs distributed almost evenly across the star forming main sequence with little bias in distribution with mass. This hints at evolution with stage.

Applying KS- or AD- tests these, however, reveals values close to 1 and therefore it is very likely that these distributions have been drawn from the same parent sample. However, due to the low counts of AGN and SFGs in our sample, using KS- and AD-tests to verify this is not optimal. Therefore, we instead investigate the global AGN fractions in our sample across different stages. We apply a weighting, as in our SFR and stellar mass distribution examples, to account for the various counts across the different stages. We then take a second weighting of the measured AGN fraction from the total size of the different samples we have. Thus, we have assumed that we have equivalent counts in each mass bin of our sample as well as approximated having the same sample size in each stage bin.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/agn-stage-dist.pdf}
    \caption{The distribution of AGN through stage with SFR and stellar mass. We do not find any major changes and evolution with AGN and SFG position through interaction stage. }
    \label{fig:agn-stage}
\end{figure}

\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
         & Stage 1 & Stage 2 & Stage 3 & Stage 4 \\
         \hline
        SFG & 135 & 212 & 337 & 199 \\
        AGN & 103 & 117 & 163 & 95
    \end{tabular}
    \caption{Breakdown in number of classified AGN and SFGs per stage. With our counts so low in this sample, it is difficult to make concrete conclusions about the evolution of AGN during interaction.}
    \label{tab:agn-sfg-breakdown}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/agn-frac-time.pdf}
    \caption{The change in AGN fraction with stage. We define the AGN fraction as the ratio of number of confirmed AGN divided by the number confirmed AGN and SFGs in the sample. Errors on each point are found by bootstrapping. We do not include any sources of which we have no information. We find a sudden drop in AGN fraction from stage 1 to stage 2, followed by a rapid increase from stage 3 to 4.}
    \label{fig:agn-frac-time}
\end{figure}

Figure \ref{fig:agn-frac-time} shows the changing AGN fraction with interaction stage. We find that the AGN fraction generally increases with interaction stage. There is a large increase from stage 1 to 2, followed by the fraction mostly remaining unchanged until stage 3. Then, in stage 4 we measure another large jump in the AGN fraction. However, before we discuss this result further, we must point out that the large error bars on our measurements and the low number counts of confirmed AGN and SFG make this difficult to interpret. We measure the errors on our sample from bootstrapping the underlying low-count population of AGN and SFGs and comparing to the change in fraction measured. Thus, while our inital values show a general increase, because of the large error bars, there could be no increase at all with stage to even a slight decline.
